# Reproducible Research: Peer Assessment 1
## Loading and preprocessing the data
### The file "activity.csv" is downloaded to the home Documents Folder
```{r}
setwd("~/Documents")
Activity<-read.csv("activity.csv",header=TRUE,comment.char = "",colClasses=
                         c("integer","character","integer"),na.strings='NA')
Activity$date<-as.Date(Activity$date,"%Y-%m-%d")
```
## What is mean total number of steps taken per day?
### Histogram of steps per day
```{r}
library(reshape2)
ActivityMelt<-melt(Activity,id.vars=c("date","interval"),na.rm=TRUE)
ActivityTotal<-dcast(ActivityMelt,date~variable,fun=sum)
hist(ActivityTotal$steps,main="Total Number of Steps Taken Each Day",
     xlab="Number of Steps")
```

### Mean and Median Steps Each Day
```{r,results="asis"}
ActivityDateMean<-dcast(ActivityMelt,date~variable,fun=mean)
ActivityDateMedian<-dcast(ActivityMelt,date~variable,fun=median,fill=0)
Report<-cbind(ActivityDateMean,ActivityDateMedian$steps)
colnames(Report)<-c("date","mean steps","median steps")
library(xtable)
print(xtable(Report),type="html")
```
## What is the average daily activity pattern?
### Time Series of Average Steps taken by time interval
```{r}
ActivityIntervalMean<-dcast(ActivityMelt,interval~variable,fun=mean)
plot(ActivityIntervalMean$interval,ActivityIntervalMean$steps,type="l",
     main="Average Daily Steps By Time Interval",xlab="Time Interval",
     ylab="Number of Steps")
``` 

### Which 5-minute interval contains the maximum number of steps?
```{r}
ActivityIntervalMean[ActivityIntervalMean$steps==
                             max(ActivityIntervalMean$steps),1]
``` 
## Imputing missing values
###Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(Activity$steps))
```
### Replace NA's with mean of interval
```{r}
Activity$steps2<-sapply(seq_len(nrow(Activity)),function (x) 
        ifelse(is.na(Activity$steps[[x]]),
               ActivityIntervalMean[ActivityIntervalMean$interval==
                                            Activity$interval[[x]],2],
               Activity$steps[[x]]))
```
### Histogram of steps per day
```{r}
ActivityMelt<-melt(Activity,id.vars=c("date","interval"),na.rm=FALSE)
ActivityTotal<-dcast(ActivityMelt,date~variable,fun=sum)
par(mfrow=c(1,2))
hist(ActivityTotal$steps,main="Total Number of Steps Taken Each Day",
     sub="Ignore NA",ylim=c(0,37),
     xlab="Number of Steps")
hist(ActivityTotal$steps2,main="Total Number of Steps Taken Each Day",
     sub="Estimate NA based on Mean for That Interval",ylim=c(0,37),
     xlab="Number of Steps")
```

### Mean and Median Steps Each Day
```{r,results="asis"}
ActivityDateMean<-dcast(ActivityMelt,date~variable,fun=mean)
ActivityDateMedian<-dcast(ActivityMelt,date~variable,fun=median,fill=0)
Report<-cbind(ActivityDateMean,ActivityDateMedian[,2:3])
colnames(Report)<-c("date","mean steps - ignore NA",
                    "mean steps - estimate NA based on mean for the interval",
                    "median steps - ignore NA",
                    "median steps - estimate NA based on mean for the interval")
print(xtable(Report),type="html")
```
### The NA values are for full days.  The result of replacing the NA values with 
### the mean for each interval is to simply create more incidences of an average
### day.  An average day has 10766.19 steps.
## Are there differences in activity patterns between weekdays and weekends?
```{r}
Activity$weekend<-ifelse(weekdays(Activity$date)=="Saturday"|
                                 weekdays(Activity$date)=="Sunday",
                         "weekend","weekday")
Activity$weekend<-as.factor(Activity$weekend)
ActivityMelt<-melt(Activity,id.vars=c("date","interval","weekend"),na.rm=TRUE)
ActivityIntervalMean<-dcast(ActivityMelt,interval+weekend~variable,fun=mean)
library(lattice)
xyplot(steps~interval|weekend,data=ActivityIntervalMean, 
       main="Average Daily Steps By Time Interval",layout=c(1,2),type="l")
```
